---
title: "Bellabeat CaseStudy"
author: "Saleh Shuaib"
date: "7/6/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```
{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
# Intorduction 

This is a  google certificate capstone analysis for bellabeat, a high-tech company that manufactures health-focused smart products . We want to find insights and trends on how people use non-bellabeat health devices. Once we find these insgits we wil apply it on bellabeat products. This will help stakeholders make data driven decisions on how to makret thier product


##Installing packages and libraries we need 
```{r}
install.packages("lmtest", repos = "http://cran.us.r-project.org")
install.packages('tidyverse')
library(tidyverse)
library(dplyr)
install.packages("lubridate")
library(lubridate)
```

##daily_activity

```{r}
daily_activity <- read.csv('dailyActivity_merged.csv')
head(daily_activity)
colnames(daily_activity)
glimpse(daily_activity)
summary(daily_activity)
```


##hourly_calories
```{r}
hourly_calories <- read.csv('hourlyCalories_merged.csv')
head(hourly_calories)
summary(hourly_calories)
```
##hourly_intensities
```{r}
hourly_intensities <- read.csv('hourlyIntensities_merged.csv')
head(hourly_intensities)
colnames(hourly_intensities)
glimpse(hourly_intensities)
summary(hourly_intensities)
```

##hourly-steps 
```{r}
hourly_steps <- read.csv('hourlySteps_merged.csv')
head(hourly_steps)
colnames(hourly_steps)
summary(hourly_steps)
```

##minuteSteps
```{r}
minute_steps <- read.csv('minuteStepsNarrow_merged.csv')
minute_steps_1 <- minute_steps %>% 
separate(ActivityMinute, into = c("Date", "Time"), sep = " ", remove = FALSE) %>%
  mutate(Date = lubridate::as_date(Date, format = "%m/%d/%Y"),
         Time = hms::as_hms(str_c(Time, ":00")))

minute_steps_1$day <- weekdays(as.Date(minute_steps_1$Date))
summary(minute_steps_1)
```

#Analysis

## How many unique member Id's are there in each of the data frames 
```{r}
n_distinct(hourly_intensities$Id)
n_distinct(hourly_calories$Id)
n_distinct(hourly_steps$Id)
```


## Merging hourly_calories, hourly_intensities and hourly_steps into one data frame. 
 we will call the new merged data frame 'hourly_merge)
```{r}
merge_1 <- merge(hourly_calories, hourly_intensities, by=c("Id","ActivityHour"))
head(merge_1)
glimpse(merge_1)
merge_2 <-  merge(merge_1, hourly_steps, by=c("Id","ActivityHour"))
head(merge_2)
glimpse(merge_2)
```

## Splitting ActivityHour column into two columns; Date & Time 
```{r}
hourly_merge <- merge_2 %>% 
  separate(ActivityHour, into = c("Date", "Time"), sep = " ", remove = FALSE) %>%
  mutate(Date = lubridate::as_date(Date, format = "%m/%d/%Y"),
         Time = hms::as_hms(str_c(Time, ":00")))

head(hourly_merge)
glimpse(hourly_merge)
head(hourly_merge)
summary(hourly_merge)
```
## Sorting our data frame by Date and Time 
```{r}
hourly_merge_1 <- hourly_merge%>% arrange(hourly_merge$Date)
hourly_merge_2 <- hourly_merge_1 %>% arrange(hourly_merge_1$Time)
head(hourly_merge_2)
View(hourly_merge_2)
```

## saving our data frame into csv  file to export it to Tableau
```{r}
write.table(hourly_merge_2, "hourly_merge_final.csv", row.names=FALSE, sep=",")
```

## Grouping and aggregating the data 
```{r}
df <- hourly_merge_2 %>% group_by(Id)
head(df)
sapply(df, class)
```

```{r}
df1 <- df                                             
df1$Id <- as.character(df1$Id) #changing Id column from integer to character
sapply(df1, class)
```

```{r}
df2 <- df1 %>% group_by(Id) %>% summarise(Value = max(TotalIntensity))  
colnames(df2)
View(df2)
sapply(df2, class)
glimpse(df2)
```
```{r}
require(data.table)
df1 <- as.data.table(df)
df3 <- df1[df1[, .I[StepTotal == max(StepTotal)], by=Id]$V1]
head(df3)
View(df3)
sapply(df3, class)
colnames(df3)
glimpse(df3)
max(df3$Calories)
df3[which.max(df3$Calories),]
df3[which.min(df3$Calories),]
summary(df3)
write.csv(x=df3, file="vizzitable")##saving df3 into csv to export into tableau
```

## plotting for data exploration 
```{r}
ggplot(data=daily_activity, aes(x=TotalSteps, y = Calories))+ geom_point() + stat_smooth(method=lm)
ggplot(data=df3, aes(x=StepTotal, y = Calories)) + geom_point()
```

